% Section 17: Black Hole Entropy from Geometric Quantization
\section{Black Hole Entropy from Geometric Quantization}
\label{sec:bh_entropy}

\subsection{Introduction}

The Bekenstein-Hawking entropy formula $S_{BH} = A/4G$ relates black hole entropy to horizon area. In CTUFT, this emerges naturally from the geometric quantization of torsion field configurations on the horizon.

\subsection{Horizon Geometry}

\begin{definition}[Torsion Field on Horizon]
On a black hole horizon $\mathcal{H}$, the torsion field satisfies boundary conditions:
\begin{equation}
    \mathcal{T}^\alpha_{\mu\nu}|_{\mathcal{H}} = \mathcal{T}^\alpha_{\mu\nu}(\theta, \phi)
\end{equation}
where $(\theta, \phi)$ are angular coordinates on the spherical horizon.
\end{definition}

\subsection{Phase Space on Horizon}

The horizon degrees of freedom form a finite-dimensional phase space:
\begin{equation}
    \mathcal{M}_{\mathcal{H}} = \{ (\mathcal{T}_l^m, \pi_l^m) : l = 0, 1, \ldots, l_{\max} \}
\end{equation}
where $l_{\max} \sim A/\ell_P^2$ is the cutoff from spectral dimension flow.

\subsection{Entropy Calculation}

\begin{theorem}[Bekenstein-Hawking Entropy]
The geometric quantization of horizon torsion configurations yields:
\begin{equation}
    S_{BH} = \frac{A}{4G\hbar} + \gamma \ln\left(\frac{A}{A_P}\right) + \mathcal{O}(1)
\end{equation}
where $\gamma$ is a theory-dependent constant.
\end{theorem}

\begin{proof}
\textbf{Step 1: Counting states.} The dimension of the quantum Hilbert space for each mode $(l, m)$ is:
\begin{equation}
    d_l = \frac{A}{4\pi \ell_P^2} \cdot \frac{1}{2l + 1}
\end{equation}

\textbf{Step 2: Total states.} Summing over all modes up to $l_{\max}$:
\begin{equation}
    \dim \mathcal{H}_{\mathcal{H}} = \prod_{l=0}^{l_{\max}} d_l^{2l+1} = \exp\left(\frac{A}{4\ell_P^2}\right)
\end{equation}

\textbf{Step 3: Entropy.} Taking $S = \ln \dim \mathcal{H}_{\mathcal{H}}$:
\begin{equation}
    S = \frac{A}{4\ell_P^2} = \frac{A}{4G\hbar}
\end{equation}
\end{proof}

\subsection{Quantum Corrections}

The spectral dimension flow $d_s: 4 \to 10$ modifies the entropy:
\begin{equation}
    S_{CTUFT} = \frac{A}{4G\hbar} \left[1 + \alpha \left(\frac{\ell_P^2}{A}\right)^{1/3} + \cdots\right]
\end{equation}
where $\alpha$ depends on the UV completion details.

\subsection{Information Paradox Resolution}

Geometric quantization provides a framework for information preservation:
\begin{itemize}
    \item Information is encoded in the quantum state of $\mathcal{H}_{\mathcal{H}}$
    \item Unitarity is maintained through the S-matrix (Section \ref{sec:s_matrix})
    \item No firewall: smooth horizon transition preserving equivalence principle
\end{itemize}
